---
title: "GeoSpatialDataAnalysis"
authors: "Chance Bradford, Yuuki Matsunari"
format: html
editor: visual
---

# Geospatial Analysis of Frequent Locations and Movement Patterns

Term project for Cpts 475 Data Science

### Load necessary libraries and Dataset

```{r}
# Load required libraries
library(dplyr)
library(tidyr)


locations = read.csv('../data/all_locations.csv', header = TRUE, sep = ",")

head(locations)

```

**extract datetime element from data and create new table variables with extracted data. removed original datetime for extracted date and time data, and the first column 'instance' data.**

```{r}

# Add time components to the existing dataframe
locations <- locations %>%
  mutate(
    datetime_posix = mdy_hm(datetime),
    date = as.Date(datetime_posix),
    time = format(datetime_posix, "%H:%M"),
    hour = hour(datetime_posix),
    minute = minute(datetime_posix)
  )

# Reorder columns to put date/time columns after the original datetime
locations <- locations %>%
  select(acc, lat, long, date, time, hour, minute)

# Display head of the modified dataframe
head(locations)

```

### Calculate the total time spent at each key location

```{r}

```

### Identify the top 5 locations where the individual spends the most time each month

```{r}

```
